Co-moving stars in \textsl{Gaia DR1}

### 2016-11-09 paper/ms.tex
work on finishing results; filled in intro/conclusion

% define macros for text
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\rave}{\project{\acronym{RAVE}}}
\newcommand{\apogee}{\project{\acronym{APOGEE}}}
\newcommand{\documentname}{\textsl{Article}}
\newcommand{\sectionname}{Section}
\newcommand{\figname}{Figure}
\newcommand{\eqname}{Equation}
\newcommand{\dr}{\acronym{DR1}}
\newcommand{\tgas}{\acronym{TGAS}}

@@

\begin{abstract}
% Context
The primary sample of the \gaia\ \textsl{First Data Release} is the brand-new
\textsl{Tycho-Gaia Astrometric Solution}. The precision and size of this sample
creates an opportunity to find new binary stars and moving groups, especially
rare binaries or sparsely populated groups.
% Aims
Here we seize this opportunity.
% Methods
We use a justified marginalized likelihood ratio test to separate
pairs of stars with surprisingly similar three-space velocities from
those consistent with being drawn independently from the field
population.  Although we perform some visualizations using a
(bias-corrected) inverse parallax as a distance, the likelihood test
works in the observable space and uses the \gaia\ noise model
responsibly.
% Results
We find pairs of co-moving stars out to very wide
separations, including separations of a few parsecs! There does not
seem to be any tidal-disruption feature in the separation distribution
at sub-parsec scales. Pairs at separations beyond a
parsec---whether they are bound in a multiple system or drifting
apart---must be very short-lived. The photometric properties of the
members of these pairs are consistent with youth. The prospects for
testing stellar models and chemical abundance measurements are
superficially discussed.
\end{abstract}

@@

Two stars moving with similar full-space velocity vectors (``co-moving stars'')
are either members of a widely-separated stellar multiplet, or remnants of a
dispersing stellar system.

Widely-separated binary stars are weakly bound binary star systems with
semi-major axes larger than $a \gtrsim 10^{-3}~{\rm pc}$. It is generally
assumed that the stars in a wide binary form from gas with the same chemical
composition and are separated enough that there is no interaction or minimal
cross-pollution as the more massive star evolves (\citealt{todo}). Because the
binding energy of these systems is so low, if wide binary stars survive for many
Gigayears their existence alone places limits on the tidal field of the Galaxy
(\citealt{todo}) and the graininess of the gravitational field (\citealt{todo}).
In this picture, a remaining puzzle is how these systems survive the dense
stellar environments of the birth clusters in which they form. Wide binaries are
therefore important tools for studying small-scale star formation and the
large-scale Galactic mass distribution. These stars will appear to be co-moving
with velocity differences comparable to the binary orbital velocity at the
given separation.

At the largest separations, $a \gtrsim 1~{\rm pc}$, co-moving stars are likely
either remnants of disrupting stellar systems or young wide binary stars: at
these extreme separations, the Galactic tidal field would disrupt binary stars
within $t_{\rm disrupt} \lesssim XX~{\rm Gyr}$. Unbound open clusters with
velocity dispersions $\sigma_v \lesssim 1~\kms$ in the Galactic disk disrupt
quickly leaving small streams of stars with the same chemical abundances
(\citealt{chemicaltaggingstuff}). Disrupting or disrupted globular clusters and
dwarf galaxies also produce streams of stars with similar chemical properties
but with larger velocity differences (\citealt{streampeeps}). Stars from such
systems at similar orbital phases will appear to be co-moving with velocity
differences comparable to the internal velocity dispersions, $\sigma_v \approx
1$--$10~\kms$.

To date, thousands of candidate co-moving star pairs have been identified by
searching for stars with common proper motions (\citealt{Luyten:1979,
Poveda:1994, Allen:2000, Gould:2003, Chaname:2004, Lepine:2007,
Alonso-Floriano:2015}). \todo{how is our search different? (1) we have distance, (2) bigger volume than Hipparcos, (3) we use the data properly, (4) we don't cut on distance or on magnitude of proper motion -- we cut on parallax S/N}
% Cite:
% Tokovinin & Lepine 2012
% https://ui.adsabs.harvard.edu/#abs/2012ApJ...757..170A/abstract
% https://ui.adsabs.harvard.edu/#abs/2012AJ....144..102T/abstract
% https://ui.adsabs.harvard.edu/#abs/2014ApJ...790..158A/abstract

@@

% Two stars moving with full-space velocity vector differences $\lesssim 1~\kms$
% with orbital semi-major axes $a \gtrsim 10^-3~{\rm pc}$.

Useful for:
- star cluster disruption and dispersal
- chemical abundance comparison (no interaction)
- dynamics

In either case, co-moving stars are of great interest
for studying the destruction of star clusters and dwarf galaxies, the tidal
field and small-scale properties of the mass distribution in the Milky Way, and
chemical uniformity and star formation processes.

@@

The primary data set used in this \documentname\ is the Tycho-Gaia Astrometric
Solution (\tgas), released as a part of Data Release 1 (\dr) of the Gaia mission
\citep{2016arXiv160904172G,2016arXiv160904303L}.
\tgas\ contains astrometric measurements (sky position,
parallax, and proper motions) and associated covariance matrices for a large
fraction of the \project{Tycho-2} catalog \citep{2000A&A...355L..27H} with median
astrometric precision comparable to that of the \project{Hipparcos} catalog
\citep[$\approx 0.3~{\rm mas}$;][]{2007ASSL..350.....V}. In terms of parallax
signal-to-noise ($\snr = \varpi/\sigma_\varpi$), the \tgas\ catalog contains
42385 high-precision stars with $\snr > 32$.

We construct a sample of candidate pair list as follows.
We first apply a global parallax S/N cut of 8 to \tgas, which leaves 619,618 stars.
Then, for each star, we search a potential co-moving partner star with a
difference in tangential velocity $\Delta v_t < 10$~\kms\ within 10 pc
using RA, DEC, a point-estimate of distance, and the proper motions.
Throughout this \documentname\, we use the following estimator for
the point-estimate of distance applying a correction for the Lutz-Kelker bias
(\citealt{lutzkelker}):
\begin{equation}
  d = 1000 \, \left[\frac{\varpi}{2} \,
    \left(1 + \sqrt{1 - \frac{16}{\snr^2}} \right) \right]^{-1} \, {\rm pc}
\end{equation}
where $\varpi$ is the parallax in mas.
The difference in tangential velocity between two stars is, then,
\begin{equation}
  |\Delta \vec{v}_{\rm t}| = |d_1 \vec\mu_1 - d_2 \vec\mu_2|
\end{equation}
where $\vec\mu = (\mu_\alpha, \mu_\delta)$.

@@

  \item  However, it is not only (wide) binaries which are gravitationally
  bound that are expected to appear as co-moving.
  OB associations, moving groups, and open clusters are also of consideration.
  Since the basic unit of our choice is a pair of stars, we can expect that these astrophysical objects may be detected as networks of pairs.
  As the pair separation increases, the nature of co-moving pairs
  will change from binaries to those related to these larger objects,
  which will tend to subtend a larger angle in sky.
  Under the assumption of having the same 3D velocity,
  two stars of a co-moving pair can essentially be thought of as projections
  of this velocity onto sphere at different viewing angles.
  The larger the difference in viewing angles is, the larger the difference in tangential 
  velocities will be.
  Due to this projection effect, a population of genuine co-moving pairs
  will extend to larger $\Delta v_t$ at larger separation.
  This indeed can be seen in Figure~\ref{fig:dv-sep} as an over-density
  in lower right corner that gets thinner as $\Delta v_t$ increases.

@@

\begin{figure*}[p]
  \begin{center}
    \includegraphics[width=\textwidth]{figures/sep_dvtan.pdf}
  \end{center}
  \caption{%
    Difference in tangential velocity vs. physical separation for candidate
    pairs of stars with separation $< 10$~pc and $|\Delta v_t| < 10$~\kms.
    \label{fig:dv-sep}}
\end{figure*}

@@

The abundance of pairs of stars with small velocity difference in
\figname~\ref{fig:dv-sep} suggests that there are a
significant number of co-moving stars in the \tgas\ data at a range
of separations.
Here we develop a method to select high-confidence co-moving
stars that properly incorporates the complex uncertainties associated with the
\gaia\ data. We make the following assumptions in order to construct a
statistical model (a likelihood function with explicit priors on our
parameters):
\begin{itemize}
  \item We assume that the uncertainties in the data---parallax, $\varpi$, and
    two proper motion components, $\propm = (\begin{array}[t]{c c} \mu_\alpha &
    \mu_\delta\end{array})^\mathsf{T}$---are Gaussian with known covariances
    $\mat{C}$.
    \todo{SMOH: mention GAIA paper that says Gaussian is good approximation?}
    % http://adsabs.harvard.edu/abs/2012A&A...538A..78L
  \item We assume that the 3-space velocities of a given pair of stars
    $(\vec{v}_i, \vec{v}_j)$ in the \tgas\ sample (relative to the solar system
    barycenter) are either (1) co-moving with velocity drawn from a velocity
    prior $p(\vec{v})$ and  velocity difference drawn from a zero-mean Gaussian
    with velocity dispersion $s$, or (2) individually drawn from a velocity
    prior $p(\vec{v})$.
\end{itemize}

@@

We compute the fully marginalized likelihood (FML) that a given pair of stars
has the same 3-space velocity with a small velocity difference drawn from a
zero-mean Gaussian with variance $s^2$ (hypothesis 1, $\mathcal{L}_1$), and
the FML of the stars having different 3-space velocities (hypothesis 2,
$\mathcal{L}_2$).
We use the FML ratio $\mathcal{L}_1/\mathcal{L}_2$ as a scalar for selecting
candidate co-moving stars, as described below in more detail.
To compute these FMLs, the likelihood functions $L_i, L_j$ are marginalized
over true 3-space velocity and distance for each star in the pair $(i,j)$.
\begin{align}
  \mathcal{L}_1 &=
    \int \, \dd d_i \, \dd d_j \, \dd^3 \vec{v} \,
    L_i(\vec{v}, d_i, s^2) \,
    L_j(\vec{v}, d_j, s^2) \,
    p(\vec{v}) \, p(d_i \given \varpi_i) \, p(d_j \given \varpi_j) \\
  \mathcal{L}_2 &=
    \int \, \dd d_i \, \dd d_j \, \dd^3 \vec{v}_i \, \dd^3 \vec{v}_j \,
    L_i(\vec{v}_i, d_i, 0) \,
    L_j(\vec{v}_j, d_j, 0) \,
    p(\vec{v}_i) \, p(\vec{v}_j) \, p(d_i \given \varpi_i) \, p(d_j \given \varpi_j). \label{eq:hyp2}
\end{align}
The marginalization integral for hypothesis 2 can be split into the product of
two simpler integrals $\mathcal{L}_2 = Q_i \, Q_j$ where
\begin{equation}
  Q = \int \, \dd d \, \dd^3 \vec{v} \, L(\vec{v}, d, 0) \, p(\vec{v}) \, p(d\given\varpi)
\end{equation}
If the velocity prior $p(\vec{v})$ is also Gaussian, the integrals over velocity
in both cases can be performed analytically:
We use a mixture of three isotropic, zero-mean Gaussian distributions
\begin{equation}
  p(\vec{v}) = \sum_{m=1}^3 \, w_m \mathcal{N}(0, \sigma_{v,m}^2)
\end{equation}
with velocity dispersions $(\sigma_{v,1}, \sigma_{v,2}, \sigma_{v,3}) = (15, 30, 50)
~\kms$ and weights $(w_1,w_2,w_3) = (0.3, 0.55, 0.15)$
meant to encompass young thin disk stars to halo stars.
We derive the relevant expressions in Appendix~\ref{sec:appendix}.
After marginalizing over velocity, the likelihood integrands only depend on
distance; we numerically compute the integrals over the true distances to each
star in a pair, $d_1,d_2$, using Monte Carlo integration with $K$ samples from
the distance posterior pdfs:
\begin{equation}
  \int \, \dd d \, \tilde{L}(d) \, p(d\given\varpi) \approx
    \frac{1}{K} \, \sum_k^K \, \tilde{L}(d_k)
\end{equation}
where $\tilde{L}(d)$ is the velocity-marginalized likelihood function.
Through experimentation, we have found that $K=128$ samples are sufficient for
estimating the above integrals for stars with a wide range in parallax
signal-to-noise. \todo{(SMOH: still unsure about this)}

\todo{discuss vorbit}

\todo{APW/SMOH: in results, specify what we set $s^2$ to:}
To search for wide binaries, we set $s^2 = \frac{2 \, G \, \msun}{|\vec{x}_i-\vec{x}_j|}$.
To search for dissolving open clusters, we set $s^2 = 1~(\kms)^2$.

@@

This section is divided into three parts. First, we discuss and justify a cut of
the likelihood ratio to select candidate co-moving pairs.
Second, we present the statistics and properties of co-moving pairs, and discuss its nature.
Finally, we describe our catalog.

\subsection{Selecting candidate co-moving pairs}

@@

\begin{figure*}[p]
  \begin{center}
    \includegraphics[width=\textwidth]{figures/likelihoodratios.pdf}
  \end{center}
  \caption{%
    Left: Density histogram of likelihood ratios for candidate pairs and random pairs.
    Along with the distribution for the entire candidate pairs (black thick line),
    we show the distribution for subsets in slices of $\Delta v_t$.
    Right: $\Delta v_t$ vs separation in slices of $\ln \mathcal{L}_1 /\mathcal{L}_2$.
    \label{fig:likelihoodratios}}
\end{figure*}

Figure~\ref{fig:likelihoodratios} shows the likelihood ratios for all
271k candidate pairs examined.
As discussed in \S~\ref{sec:data}, we expect a correlation between likelihood
ratios of pairs, and their distribution in $\Delta v_t$ vs separation plane.
Specifically, as we sweep through from small $\Delta v_t$ to large, we expect
the population of pairs to change from genuinely co-moving to random.
This becomes clear when we look at the distribution of likelihood ratios for
slices of $\Delta v_t$ at 0.5, 5, and 9.5~\kms.
Pairs with $|\Delta v_t-0.5|<0.5$ are most likely actual co-moving pairs, and
their likelihood distribution is narrowly peaked above 6.
The distribution peaks at lower values and gets broader as $\Delta v_t$ increases,
and the number of random pairs increasingly dominate.
On the right panel of Figure~\ref{fig:likelihoodratios}, we show
how the distribution of pairs on $\Delta v_t$ vs separation plane changes
with high to low $\ln \mathcal{L}_1 /\mathcal{L}_2$ ratios.
This is in agreement with our discussion in \S~\ref{sec:data}.
Finally, as a test,
we compute the likelihood ratios for 200,000 random pairs of stars with the same
parallax signal-to-noise ratio cut as candidate pairs ($S/N > 8$).
Shown as gray shaded histogram, this distribution peaks at lower value ($\approx 2$),
and is clearly separated from highly probable co-moving pairs.

@@

We present the statistics of the resulting co-moving pairs.
We also match the pairs with known open clusters in the Milky Way to
provide information on cluster membership.

Once we have decided the co-moving pairs from candidate pair list,
we can count how many co-moving neighbors a star has.
This is presented in Figure~\ref{fig:hist_Nneighbors}.
It is clear from this figure that we have many aggregates of co-moving stars, most
likely an open cluster or a moving group. We also see that there is still
a large number of stars which only one or two co-moving neighbors.
In order to select pairs that are not associated with any other pair,
we separate a subset of co-moving pairs whose member is a mutually exclusive
co-moving neighbor to the other member (``mutually exclusive pairs'').
The number of mutually exclusive pairs is listed on the last column of
Table~\ref{tab:candidatecounts}.

We now examine the separation distribution of co-moving pairs in
Figure~\ref{fig:hist_separation}.

Finally, we show the distribution co-moving pairs in galactic longitude and
distance in Figure~\ref{fig:glon_d_pairlines}.
In the top panel, we show all co-moving pairs with contamination rate of $0.01~\%$
(undersampled by a factor of 4 in order to unclutter the plot) as connected line segments.
We find many aggregates of co-moving pairs, which coincides with the positions of
known Milky Way star clusters \citep{Kharchenko:2016aa} indicated by red circles.
Clumps of co-moving pairs at $l=300-360$~degrees and $d=100-200$~pc correspond
to the location of OB associations
Upper Scorpius, Upper Centaurus Lupus, and Lower Centaurus Crux
\citep{de-Zeeuw:1999aa}.
The strong correlation with known co-moving structures disappears when we only
consider mutually exclusive pairs (bottom panel).

\begin{figure*}[p]
  \begin{center}
    \includegraphics[width=\textwidth]{figures/dist_networksize.pdf}
  \end{center}
  \caption{%
    Histogram of co-moving network sizes.
    \label{fig:hist_Nneighbors}}
\end{figure*}

\begin{figure*}[p]
  \begin{center}
    \includegraphics[width=\textwidth]{figures/hist_separation.pdf}
    \includegraphics[width=\textwidth]{figures/hist_separation_mutexc.pdf}
  \end{center}
  \caption{%
    Top: Separation distribution of co-moving pairs of stars for varying degrees
    of contamination due to random pairs. The distribution peaks around 10~pc,
    which is similar to the typical tidal radius of clusters in the Milky Way
    \citep{Kharchenko:2013aa}.
    Bottom: Same as the previous plot but only for mutually exclusive pairs.
    \label{fig:hist_separation}}
\end{figure*}

\begin{figure*}[p]
  \begin{center}
    \includegraphics[width=\textwidth]{figures/glon_d_pie.pdf}
  \end{center}
  \caption{%
    Panoramic view of co-moving pairs of stars. The angle is the Galactic
    longitude, and the distance is in pc.
    Pair networks of size 2 and 3 are connected by gray and blue lines.
    Pair networks of size 4 and larger are plotted with a unique color.
    Black circles mark the positions of known Milky Way star clusters
    from \citet{Kharchenko:2016aa}.
    \label{fig:glon_d_pairlines}}
\end{figure*}

>>>>

% define macros for text
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\rave}{\project{\acronym{RAVE}}}
\newcommand{\apogee}{\project{\acronym{APOGEE}}}
\newcommand{\tmass}{\project{\acronym{2MASS}}}
\newcommand{\documentname}{\textsl{Article}}
\newcommand{\sectionname}{Section}
\newcommand{\figname}{Figure}
\newcommand{\eqname}{Equation}
\newcommand{\dr}{\acronym{DR1}}
\newcommand{\tgas}{\acronym{TGAS}}

@@

\begin{abstract}
% Context
The primary sample of the \gaia\ \textsl{First Data Release} is the brand-new
\textsl{Tycho-Gaia Astrometric Solution}. The precision and size of this sample
creates an opportunity to find new binary stars and moving groups, especially
rare binaries or sparsely populated groups.
% Aims
Here we seize this opportunity.
% Methods
We use a justified marginalized likelihood ratio test to separate
pairs of stars with surprisingly similar three-space velocities from
those consistent with being drawn independently from the field
population.  Although we perform some visualizations using a
(bias-corrected) inverse parallax as a distance, the likelihood test
works in the observable space and uses the \gaia\ noise model
responsibly.
% Results
We find pairs of co-moving stars out to very wide
separations up to 10~pc, the limit of our search.
Some of these pairs form larger networks through mutual co-moving neighbors,
many of which correspond to the known open clusters, and stars in the OB associations.
However, there still is a huge number of very wide ($>3$~pc) separation pairs of
stars that are co-moving.
We discuss their nature, and the prospects for testing stellar models and chemical
abundance measurements using our catalog.

% HOGG:
% There does not seem to be any tidal-disruption feature in the separation distribution
% at sub-parsec scales. Pairs at separations beyond a
% parsec---whether they are bound in a multiple system or drifting
% apart---must be very short-lived. The photometric properties of the
% members of these pairs are consistent with youth. The prospects for
% testing stellar models and chemical abundance measurements are
% superficially discussed.
\end{abstract}

@@

Stars that are roughly co-located, and moving with similar space velocity (``co-moving stars'')
are of special interest in many branches of astrophysics.

At small scales ($0.001-1$~pc), they correspond to wide binaries (and triples, or
multiples) that are
either weakly gravitationally bound or currently being dispersed.
Because they have low binding energies,
a sample of wide binaries is valuable for investigating
the Galactic dynamical environment.
These systems must have survived their dynamic birth environment which
provides a window to the star formation process.
They may also be disrupted by Galactic tides, or other massive objects
such as other stars, molecular clouds, or MACHOs \citep{Yoo:2004aa,Allen:2014aa}.
Wide binaries are also good testbeds for stellar evolution models.
For example, main sequence-red giant wide binaries are especially valuable
in testing consistency between stellar atmosphere models of these two types
as the two stars were born at the same time with the same chemical compositions,
but remained isolated since due to wide separation.
Finally, calibration of stellar parameters of low-mass stars (e.g., M dwarfs),
which dominate the stellar content of the Galaxy by number, can benefit from
a larger sample of widely separated binaries containing a low mass star and a
much brighter FGK star whose stellar parameters are easier to measure
\citep[e.g.,][]{Rojas-Ayala:2012aa}.

At larger scales ($\gtrsim 1$~pc), they include moving groups, associations, and star clusters.
The origin of moving groups is still under active debate: are they remnants of a coeval
star formation event with similar chemical composition? Or are they formed by
dynamical effects of non-axisymmetric features of the Galaxy such as spirals
and bars? With the recent improvements in measuring chemical abundances
using high and low resolution spectroscopy, we can now start to explore these questions
in unprecedented width (statistics) and depth (detail). 
The dynamics of cluster dissolution provides important clues to understanding
the star formation history and the dynamical evolution of the Milky Way.
In the halo, we have known the presence of disrupting globular clusters and dwarf galaxies
\citep[e.g.,][]{Ibata:1994aa}.
The tidal streams formed during this process are modeled to infer the
parameters of the Galactic potential at large scales.
Similar process is at work with star clusters in the disk. However, the dynamical
time is much shorter, and the dynamics will be much more complex due to other
perturbers in the disk.

To date, thousands of candidate co-moving star pairs have been identified by
searching for stars with common proper motions
(\citealt{Poveda:1994aa,Allen:2000aa,Chaname:2004aa,Lepine:2007aa,Alonso-Floriano:2015aa}).
Here, we use the recent first data release of \gaia\ which includes precise distances,
enabling us to ask whether two stars share the same {\it physical} velocity rather than
just the projections in the proper motion space.

%TODO: highlight the merits of this search:
% - we have distance,
% - bigger volume than Hipparcos,
% - we use the data properly, 
% - we don't cut on distance or on magnitude of proper motion -- we cut on parallax S/N}

@@

@@

The primary data set used in this \documentname\ is the Tycho-Gaia Astrometric
Solution (\tgas), released as a part of Data Release 1 (\dr) of the Gaia mission
\citep{Gaia-Collaboration:2016aa,Lindegren:2016aa}.
\tgas\ contains astrometric measurements (sky position,
parallax, and proper motions) and associated covariance matrices for a large
fraction of the \project{Tycho-2} catalog \citep{2000A&A...355L..27H} with median
astrometric precision comparable to that of the \project{Hipparcos} catalog
\citep[$\approx 0.3~{\rm mas}$;][]{2007ASSL..350.....V}. In terms of parallax
signal-to-noise ($\snr = \varpi/\sigma_\varpi$), the \tgas\ catalog contains
42385 high-precision stars with $\snr > 32$.

We construct a sample of candidate pair list as follows.
We first apply a global parallax S/N cut of 8 to \tgas, which leaves 619,618 stars.
Then, for each star, we search a potential co-moving partner star with a
difference in tangential velocity $\Delta v_t < 10$~\kms\ within 10 pc
using RA, DEC, a point-estimate of the distance, and the proper motions.
Throughout this \documentname\, we use the following estimator for
the point-estimate of the distance applying a correction for the Lutz-Kelker bias
(\citealt{Lutz:1973aa}):
\begin{equation}
  d = 1000 \, \left[\frac{\varpi}{2} \,
    \left(1 + \sqrt{1 - \frac{16}{\snr^2}} \right) \right]^{-1} \, {\rm pc}
\end{equation}
where $\varpi$ is the parallax in mas.
The difference in tangential velocity between two stars is, then,
\begin{equation}
  |\Delta \vec{v}_{\rm t}| = |d_1 \vec\mu_1 - d_2 \vec\mu_2|
\end{equation}
where $\vec\mu = (\mu_\alpha, \mu_\delta)$.

@@

  \item  However, it is not only (wide) binaries which are gravitationally
  bound that are expected to appear as co-moving.
  OB associations, moving groups, and open clusters are also of consideration.
  Since the basic unit of our choice is a pair of stars, we can expect that these astrophysical objects may be detected as networks of pairs.
  As the pair separation increases, the nature of co-moving pairs
  will change from binaries to those related to these larger objects,
  which generally subtend a larger angle in sky.
  Since the proper motions of two stars with same 3D velocity
  are essentially projections of this velocity onto sphere at
  two different viewing angles,
  the larger the difference in viewing angles is, the larger the difference in tangential 
  velocities will be.
  Due to this projection effect, a population of genuine co-moving pairs
  will extend to larger $\Delta v_t$ at larger separation.
  This indeed can be seen in Figure~\ref{fig:dv-sep} as an over-density
  in lower right corner that gets thinner as $\Delta v_t$ increases.

@@

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=\textwidth]{figures/sep_dvtan.pdf}
  \end{center}
  \caption{%
    Difference in tangential velocity vs. physical separation for candidate
    pairs of stars with separation $< 10$~pc and $|\Delta v_t| < 10$~\kms.
    \label{fig:dv-sep}}
\end{figure*}

@@

The abundance of pairs of stars with small velocity difference in
\figname~\ref{fig:dv-sep} suggests that there are a
significant number of co-moving stars in the \tgas\ data at a range
of separations.
Here we develop a method to select high-confidence co-moving
stars that properly incorporates the complex uncertainties associated with the
\gaia\ data. We make the following assumptions in order to construct a
statistical model (a likelihood function with explicit priors on our
parameters):
\begin{itemize}
  \item We assume that the uncertainties in the data---parallax, $\varpi$, and
    two proper motion components, $\propm = (\begin{array}[t]{c c} \mu_\alpha &
    \mu_\delta\end{array})^\mathsf{T}$---are Gaussian with known covariances
    $\mat{C}$. The values of covariances are provided by the \gaia\ data
    \citep{Lindegren:2012aa,Lindegren:2016aa}.
  \item We assume that the 3-space velocities of a given pair of stars
    $(\vec{v}_i, \vec{v}_j)$ in the \tgas\ sample (relative to the solar system
    barycenter) are either (1) co-moving with velocity drawn from a velocity
    prior $p(\vec{v})$ and  velocity difference drawn from a zero-mean Gaussian
    with velocity dispersion $s$, or (2) individually drawn from a velocity
    prior $p(\vec{v})$.
\end{itemize}

@@

We compute the fully marginalized likelihood (FML) that a given pair of stars
has the same 3-space velocity with a small velocity difference drawn from a
zero-mean Gaussian with variance $s^2$ (hypothesis 1, $\mathcal{L}_1$), and
the FML of the stars having independent 3-space velocities (hypothesis 2,
$\mathcal{L}_2$).
We use the FML ratio $\mathcal{L}_1/\mathcal{L}_2$ as a scalar for selecting
candidate co-moving stars, as described in $\S~\ref{sub:selection}$ in more detail.
To compute these FMLs, the likelihood functions $L_i, L_j$ are marginalized
over true 3-space velocity and distance for each star in the pair $(i,j)$.
\begin{align}
  \mathcal{L}_1 &=
    \int \, \dd d_i \, \dd d_j \, \dd^3 \vec{v} \,
    L_i(\vec{v}, d_i, s^2) \,
    L_j(\vec{v}, d_j, s^2) \,
    p(\vec{v}) \, p(d_i \given \varpi_i) \, p(d_j \given \varpi_j) \\
  \mathcal{L}_2 &=
    \int \, \dd d_i \, \dd d_j \, \dd^3 \vec{v}_i \, \dd^3 \vec{v}_j \,
    L_i(\vec{v}_i, d_i, 0) \,
    L_j(\vec{v}_j, d_j, 0) \,
    p(\vec{v}_i) \, p(\vec{v}_j) \, p(d_i \given \varpi_i) \, p(d_j \given \varpi_j). \label{eq:hyp2}
\end{align}
The marginalization integral for hypothesis 2 can be split into the product of
two simpler integrals $\mathcal{L}_2 = Q_i \, Q_j$ where
\begin{equation}
  Q = \int \, \dd d \, \dd^3 \vec{v} \, L(\vec{v}, d, 0) \, p(\vec{v}) \, p(d\given\varpi)
\end{equation}
If the velocity prior $p(\vec{v})$ is also Gaussian, the integrals over velocity
in both cases can be performed analytically:
We use a mixture of three isotropic, zero-mean Gaussian distributions
\begin{equation}
  p(\vec{v}) = \sum_{m=1}^3 \, w_m \mathcal{N}(0, \sigma_{v,m}^2)
\end{equation}
with velocity dispersions $(\sigma_{v,1}, \sigma_{v,2}, \sigma_{v,3}) = (15, 30, 50)
~\kms$ and weights $(w_1,w_2,w_3) = (0.3, 0.55, 0.15)$
meant to encompass young thin disk stars to halo stars.
We derive the relevant expressions in Appendix~\ref{sec:appendix}.
After marginalizing over velocity, the likelihood integrands only depend on
distance; we numerically compute the integrals over the true distances to each
star in a pair, $d_1,d_2$, using Monte Carlo integration with $K$ samples from
the distance posterior pdfs:
\begin{equation}
  \int \, \dd d \, \tilde{L}(d) \, p(d\given\varpi) \approx
    \frac{1}{K} \, \sum_k^K \, \tilde{L}(d_k)
\end{equation}
where $\tilde{L}(d)$ is the velocity-marginalized likelihood function.
Through experimentation, we have found that $K=128$ samples are sufficient for
estimating the above integrals for stars with a wide range in parallax
signal-to-noise.
%TODO: SMOH: still unsure about this

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=\textwidth]{figures/a_vorb.pdf}
  \end{center}
  \caption{%
    Binary orbital velocity as a function of semi-major axis for $2~M_\odot$
    binary.
    \label{fig:vorbit}}
\end{figure*}

For wide binaries, the assumption of same 3D velocity can break down
for high precision proper motion measurements because of the orbital
velocity (Figure~\ref{fig:vorbit}).
To account for this, we set $s^2 = \frac{2 \, G \, \msun}{|\vec{x}_i-\vec{x}_j|}$.

@@

This section is divided into three parts. First, we discuss and justify a cut of
the likelihood ratio to select candidate co-moving pairs.
Second, we present the statistics and properties of co-moving pairs, and discuss its nature.
Finally, we describe our catalog of candidate co-moving pairs, the main product
of this study.

\subsection{Selecting candidate co-moving pairs}
\label{sub:selection}

@@

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=\textwidth]{figures/likelihoodratios.pdf}
  \end{center}
  \caption{%
    Left: Density histogram of likelihood ratios for candidate pairs and random pairs.
    Along with the distribution for the entire candidate pairs (black thick line),
    we show the distribution for subsets in slices of $\Delta v_t$.
    Right: $\Delta v_t$ vs separation in slices of $\ln \mathcal{L}_1 /\mathcal{L}_2$.
    \label{fig:likelihoodratios}}
\end{figure*}

Figure~\ref{fig:likelihoodratios} shows the likelihood ratios for all
271k candidate pairs examined.
As discussed in \S~\ref{sec:data}, we expect a correlation between likelihood
ratios of pairs, and their distribution in $\Delta v_t$ vs separation plane.
Specifically, as we sweep through from small $\Delta v_t$ to large, we expect
the population of pairs to change from genuinely co-moving to random.
This becomes clear when we look at the distribution of likelihood ratios for
slices of $\Delta v_t$ centered at 0.5, 5, and 9.5~\kms.
Pairs with $|\Delta v_t-0.5|<0.5$ are most likely actual co-moving pairs, and
their likelihood distribution is narrowly peaked above 6.
The distribution peaks at lower values and gets broader as $\Delta v_t$ increases,
and the number of random pairs increasingly dominate.
On the right panel of Figure~\ref{fig:likelihoodratios}, we show
how the distribution of pairs on $\Delta v_t$ vs separation plane changes
with decreasing $\ln \mathcal{L}_1 /\mathcal{L}_2$ ratios.
This is in agreement with our discussion in \S~\ref{sec:data}.
Finally, as a test,
we compute the likelihood ratios for 200,000 random pairs of stars with the same
parallax signal-to-noise ratio cut as candidate pairs ($S/N > 8$).
Shown as gray shaded histogram in Figure~\ref{fig:likelihoodratios},
this distribution peaks at a much lower value ($\approx 2$),
and is clearly separated from highly probable co-moving pairs.

@@

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=\textwidth]{figures/dist_networksize.pdf}
  \end{center}
  \caption{%
    Histogram of co-moving network sizes.
    \label{fig:hist_Nneighbors}}
\end{figure*}

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=\textwidth]{figures/graphviz_examples.pdf}
  \end{center}
  \caption{%
    Visualizations of a few example networks of co-moving pairs of stars.
    Each star (node) is marked as red circles, and a line (edge) is drawn
    between two stars if they are co-moving by our selection criteria
    (see \S~\ref{sub:selection}). On left, we show the largest network
    found in this study corresponding to the Pleiades star cluster.
    \label{fig:graphviz_examples}}
\end{figure*}

\begin{figure*}[htbp]
  \begin{center}
    \includegraphics[width=\textwidth]{figures/glon_d_pie.pdf}
  \end{center}
  \caption{%
    Panoramic view of co-moving pairs of stars. The angle is the Galactic
    longitude, and the distance is in pc.
    Pair networks of sizes less 6 are connected by gray lines.
    Pair networks of size larger than 6 are plotted with a unique color per network.
    We also show the positions of known Milky Way star clusters
    from \citet{Kharchenko:2016aa} as empty purple circles,
    and stars in OB associations from \citet{de-Zeeuw:1999aa} as
    black circles.
    \label{fig:glon_d_pairlines}}
\end{figure*}

Once we have decided the co-moving pairs from the candidate pair list,
these pairs form an undirected graph where stars are nodes, and
edges between the nodes exist for co-moving pairs of stars.
A star may have multiple co-moving neighbors, and two stars may either
be directly or indirectly connected by a set of edges.
We divide the network into connected components\footnote{A connected component of an undirected graph $G$
is a subgraph of $G$ in which any two nodes are connected to each other by a path.
We use the term ``network'' loosely to refer to a graph or its connected components.},
and show the distribution of their sizes in Figure~\ref{fig:hist_Nneighbors}.
The most common are size 2 networks, which mean mutually exclusive co-moving
pairs. However, it is clear that there are many aggregates of co-moving stars
discovered by looking for co-moving pairs.
These aggregates are likely moving groups, OB associations or star clusters.
There are 4555 connected components among 10,606 unique stars.
The maximum size of connected components is 151.
We show this largest network along with four randomly selected
examples of networks with sizes greater than 4 in Figure~\ref{fig:graphviz_examples}.
The largest network corresponds to the Pleiades open cluster.
One particular star in this network, TYC 1799-272-1, has 94 directly connected
co-moving neighbors, and is located at the center of the network.
